---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# overedge

<!-- badges: start -->
<!-- badges: end -->

The goal of overedge is to provide useful functions for making maps with R. This is a collection of miscellaneous functions primarily for working with ggplot2 and sf.

## Installation

You can install the development version of overedge like so:

``` r
remotes::install_github("elipousson/overedge")
```

## Example

`overedge` currently provides a single function (`geom_sf_icon`) that wraps `ggsvg::geom_point_svg()` to provide an convenient way to make icon maps.

You can create maps using a single named icon that matches one of the icons in `map_icons`.

```{r}
library(overedge)
library(ggplot2)
library(sf)

nc <- st_read(system.file("shape/nc.shp", package="sf"))
ggplot() +
  geom_sf_icon(data = nc, icon = "point-start", size = 8) +
  theme_void()
```

You can also use an `icon` column from the provided sf object.

```{r}
nc$icon <- rep(c("1", "2", "3", "4"), nrow(nc)/4)
ggplot() +
  geom_sf_icon(data = nc, size = 5) +
  theme_void()
```

Check `map_icons` to see all supported icon names.

```{r}
head(map_icons)
```

