---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# overedge

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/overedge)](https://CRAN.R-project.org/package=overedge)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
 [![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](https://opensource.org/licenses/MIT)
[![Codecov test coverage](https://codecov.io/gh/elipousson/overedge/branch/main/graph/badge.svg)](https://app.codecov.io/gh/elipousson/overedge?branch=main)
<!-- badges: end -->

The goal of overedge is to provide useful functions for making maps with R. This is a collection of miscellaneous functions primarily for working with ggplot2 and sf.

## Installation

You can install the development version of overedge like so:

``` r
remotes::install_github("elipousson/overedge")
```

## Examples

`overedge` currently provides a variety of functions for accessing spatial data, modifying simple feature or bounding box objects, and creating or formatting maps with ggplot2.

### Make icon maps with sf objects and ggplot2

`geom_sf_icon` wraps `ggsvg::geom_point_svg()` to provide an convenient way to make icon maps.

You can create maps using a single named icon that matches one of the icons in `map_icons`.

```{r}
library(overedge)
library(ggplot2)
library(sf)

nc <- st_read(system.file("shape/nc.shp", package = "sf"))
nc <- st_transform(nc, 4326)
theme_set(theme_void())

nc_map <-
  ggplot() +
  geom_sf(data = nc)

nc_map +
  geom_sf_icon(data = nc, icon = "point-start", size = 8)
```

You can also use an `icon` column from the provided sf object.

```{r}
nc$icon <- rep(c("1", "2", "3", "4"), nrow(nc) / 4)

nc_map +
  geom_sf_icon(data = nc, size = 5)
```

Check `map_icons` to see all supported icon names.

```{r}
head(map_icons)
```

### Scale and rotate sf objects

`st_scale_rotate()` is a convenience function for apply affine transformations to sf objects.

```{r}
nc_rotated <- st_scale_rotate(nc, scale = 0.5, rotate = 15)

nc_map +
  geom_sf(data = nc_rotated, fill = NA, color = "red")
```

### Create inscribed squares in sf objects

```{r}
nc_squares <- st_inscribed_square(nc)

nc_map +
  geom_sf(data = nc_squares, fill = NA, color = "red")
```

### Add a neatline to a map

`layer_neatline()` hides major grid lines and axis label by default. The function is useful to draw a neatline around a map at a set aspect ratio.

```{r}
nc_map +
  layer_neatline(
    data = nc,
    asp = "6:4",
    color = "gray60", size = 2, linetype = "dashed"
  )
```

`layer_neatline()` can also be used to focus on a specific area of a map with the option to apply a buffer as a distance or ratio of the diagonal distance for the input data. The `label_axes` and `hide_grid` paramters will not override a set ggplot theme.

```{r}
theme_set(theme_minimal())

nc_map +
  layer_neatline(
    data = nc[1, ],
    diag_ratio = 0.5,
    asp = 1,
    color = "black",
    label_axes = "--EN",
    hide_grid = FALSE
  )
```
