---
title: "Getting remote data and photos using API functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting remote data and photos using API functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(overedge)
library(ggplot2)
```

```{r get_location, collapse=TRUE}
location <-
  get_location(
    type = "neighborhoods",
    name = "Downtown",
    package = "mapbaltimore"
  )

basemap <-
  ggplot() +
  layer_location_data(
    data = "streets",
    package = "mapbaltimore",
    location = location,
    color = "gray80"
  )
```


# Get data from an ArcGIS FeatureServer or MapServer

A large number of cities have adopted [ArcGIS Hub](https://www.esri.com/en-us/arcgis/products/arcgis-hub/overview) as an open data platform for spatial and non-spatial data. get_esri_data wraps the esri2sf package to facilitate getting spatial data by bounding box or by area name.

For example, this FeatureServer from [Open Baltimore](https://data.baltimorecity.gov/datasets/landmarks-1/explore) provides a list of designated landmarks in Downtown Baltimore.

```{r}
# https://data.baltimorecity.gov/datasets/landmarks-1/explore
url <- "https://opendata.baltimorecity.gov/egis/rest/services/Hosted/Landmarks/FeatureServer/0"

landmarks <-
  get_esri_data(
    location = location,
    url = url
  )

landmarks <-
  dplyr::select(landmarks, fid, name, address)

basemap +
  layer_location_data(
    data = landmarks
  ) +
  layer_location_data(
    data = landmarks,
    geom = "text_repel",
    size = 2
  ) +
  layer_neatline(
    data = location
  )
```

Even non-spatial Tables can be filtered to a location if the table includes latitude and longitude columns.

# Get data from Open Street Map

To get data from Open Street Map, you may need to refer to the [OSM Wiki](https://wiki.openstreetmap.org) for more information about keys and tags such as [tourism](https://wiki.openstreetmap.org/wiki/Key:tourism) and [artwork](https://wiki.openstreetmap.org/wiki/Tag:tourism%3Dartwork).

```{r, eval=FALSE}
osm_artwork <-
  get_osm_data(
    location = location,
    key = "tourism",
    value = "artwork",
    geometry = "point"
  )

basemap +
  layer_location_data(
    data = osm_artwork,
    aes(color = material)
  ) +
  layer_neatline(
    data = location,
    expand = TRUE
  ) +
  theme_legend(position = "topleft")
```

# Get data about photos from Flickr

The get_flickr_photos function depends on a [development version of the FlickrAPI package](https://github.com/elipousson/FlickrAPI) so I am not currently able to render the vignette to show


```{r, eval = FALSE}
photos_sf <-
  get_flickr_photos(
    location = location,
    img_size = "s",
    tags = "monument",
    orientation = "landscape",
    geometry = TRUE # this is the default
  )
```

The following is an experimental function using the leafpop package to make an interactive photo map based on photos from Flickr. 

```{r, eval = FALSE}
mapview_flickr <- function(photos,
                           img_size = "s") {
  img_size <-
    match.arg(
      img_size,
      c("sq", "t", "s", "q", "m", "n", "z", "c", "l", "o")
    )

  leaflet_map <-
    leaflet::leaflet() %>%
    leaflet::addTiles() %>%
    leaflet::addCircleMarkers(data = photos, group = "photos")

  img_url <- photos$img_url
  img_width <- photos$img_width
  img_height <- photos$img_height

  photo_map <-
    leafpop::addPopupImages(
      map = leaflet_map,
      image = img_url,
      width = img_width,
      height = img_height,
      group = "photos"
    )

  return(photo_map)
}


mapview_flickr(photos = photos_sf, img_size = "s")
```
